---
title: "computation time_table"
author: "Josh Alampi"
date: "2025-08-27"
output: html_document
---



# load
```{r}
#clear workspace
rm(list=ls(all=TRUE))

# load packages and functions
source(here::here("libraries.R"))
source(here::here("functions/ggplot functions.R"))

data <- read.csv(here::here("data", "raw", "QR sim study_all models_2025-09-08_1000_sims.csv")) %>% 
  select(-c(beta:true_beta)) %>%  # get rid of column we don't need
  filter(method != "bayes_adjv1") %>%  # this method does not work, do not include
  filter(model == 1) # Computation time does not differ by model. Will just use model 1 for simplicity

```


# clean
## Fix labels
```{r}
# Edit method labels
int_method_labels_cov <- c("Rank (not\ndithered)", "xy-bootstrapᵇ\n(not dithered)",
                           "Rank\n(dithered)", "xy-bootstrapᵇ\n(dithered)",
                           "Bayesianᶜ\n(unadjusted)", "Bayesianᶜ\n(adjusted)")

data <- data %>% 
  # Fix method labels
  mutate(int_method_cov = case_when(method == "riid_nodith" ~ int_method_labels_cov[1],
                                    method == "xy_nodith" ~ int_method_labels_cov[2],
                                    method == "riid_dith" ~ int_method_labels_cov[3],
                                    method == "xy_dith" ~ int_method_labels_cov[4],
                                    method == "bayes_unadj" ~ int_method_labels_cov[5],
                                    method == "bayes_adjv2" ~ int_method_labels_cov[6]
                                    )) %>% 

  ## Set the order
  mutate(int_method_cov = factor(int_method_cov, int_method_labels_cov)) %>% 
  
  # Clean up the sample size 'n' labels
  mutate(n = paste0("n = ", n)) %>% 
  
  # Clean up the tau labels
  mutate(tau = as.factor(tau))%>%
  mutate(tau = paste0("τ = ", tau))

```



# Get results, clean
For simplicity, I will only report tau = 0.5 The xy method differed slightly depending on tau, but either way it was clearly much faster than Bayesian QR, and a bit slower than rank-based QR. Better to simplify the table if the message is identical. 
```{r}
results_round <- 2

clean_data <- data %>% 
  filter(tau == "τ = 0.5") %>% 
  select(-c(tau)) %>% 
  
  # Calculation
  group_by(int_method_cov, n) %>% 
  summarise(mean_time = format(round(mean(time, na.rm = T), results_round), 
                               nsmall = results_round, trim =T),
            sd_time = format(round(sd(time, na.rm = T), results_round), 
                               nsmall = results_round, trim =T)
            ) %>% 
  
  # Clean up
  mutate(time = paste0(mean_time, " (", sd_time, ")")) %>% 
  select(-c(mean_time, sd_time)) %>% 
  
  # Make wide
  pivot_wider(names_from = c(n), 
              values_from = c(time), names_vary = "slowest") 

# Fix column names
column_names <- c("Method", 
                  "Mean (SD) computation time, seconds_n = 100",
                  "Mean (SD) computation time, seconds_n = 250",
                  "Mean (SD) computation time, seconds_n = 500",
                  "Mean (SD) computation time, seconds_n = 750"
                  )
colnames(clean_data) <- column_names # set names

```

# make flextable

## Set table labels
```{r}
title <- "Table S3. Computation timeᵃ for quantile regression methods. Model 1 and τ = 0.5 used throughout."

footer <- "a: With a Windows 10 computer, i7-7700 processor, R version 4.5.1.\nb: 1000 bootstraps used. Computation time depends on the number of bootstraps used. \nc: 10,000 Markov Chain Monte Carlo iterations used. Computation time depends on the number of iterations used."
```

## set flextable defaults
```{r}
init_flextable_defaults()

set_flextable_defaults(
  font.size = 9.5, font.family = "Calibri",
  font.color = "black",
  table.layout = "fixed",
  border.color = "black",
  line_spacing = 1,
  padding.top = 0, padding.bottom = 0,
  padding.left = 0, padding.right = 0)

```

## make flextable
```{r}
time_table <- clean_data %>%
  flextable() %>%
  separate_header(split = "_") %>% 

  # align header
  align(align = "center", part = "header") %>% # align to center by default
  align(align = "left", part = "header", j = 1) %>% 

  # align body
  align(align = "center", part = "body") %>% # align to center by default
  align(align = "left", j = c(1)) %>% # first row to the left

  # Add horizontal lines throughout
  hline(part = "body") %>%

  # Add bold lines in the header
  hline(part = "header", border = fp_border(width = 2)) %>%

  # Title and bottom caption
  add_header_lines(title, top = T) %>%
  add_footer_lines(footer) %>% 

  # set_table_properties(layout = "autofit", width = 1)
  width(j = c(1:5),
        width = c(1, rep(0.8, 4)),
        unit = "in")

time_table
```
# save
```{r}
doc <- read_docx() %>% 
  body_add_flextable(value = time_table)

base::print(doc, target = paste0("results/table_computation time", "_",  
                                 Sys.Date(), ".docx"))
```


```{r}

```











# CUT

# Time by tau
Will not use.

# Get results, clean
```{r}
results_round <- 2

clean_data_with_tau <- data %>%
  # Calculation
  group_by(int_method_cov, n, tau) %>%
  summarise(mean_time = format(round(mean(time, na.rm = T), results_round),
                               nsmall = results_round, trim =T),
            sd_time = format(round(sd(time, na.rm = T), results_round),
                               nsmall = results_round, trim =T)
            ) %>%

  # Clean up
  mutate(time = paste0(mean_time, ", (", sd_time, ")")) %>%
  select(-c(mean_time, sd_time)) %>%

  # Make wide
  pivot_wider(names_from = c(tau),
              values_from = c(time), names_vary = "slowest") %>%

  # set order
  arrange(n, int_method_cov)

# Fix column names
column_names <- c("Method", "n",
                  "Mean (SD) computation time, seconds_τ = 0.1",
                  "Mean (SD) computation time, seconds_τ = 0.5",
                  "Mean (SD) computation time, seconds_τ = 0.9"
                  )
colnames(clean_data_with_tau) <- column_names # set names

```

```{r}
time_table_with_tau <- clean_data_with_tau %>%
  flextable() %>%
  separate_header(split = "_") %>% 

  # align header
  align(align = "center", part = "header") %>% # align to center by default
  align(align = "left", part = "header", j = 1) %>% 

  # align body
  align(align = "center", part = "body") %>% # align to center by default
  align(align = "left", j = c(1)) %>% # first row to the left

  # Add horizontal lines throughout
  hline(part = "body") %>%
  
  # Add bold lines separating the different n values
  hline(i = c(6, 12, 18, 24), part = "body",
        border = fp_border(width = 2)) %>% 

  # Add bold lines in the header
  hline(part = "header", border = fp_border(width = 2)) %>%

  # Title and bottom caption
  add_header_lines(title, top = T) %>%
  add_footer_lines(footer) 

  # # set_table_properties(layout = "autofit", width = 1)
  # width(j = c(1:5),
  #       width = c(0.6, 1, 1.4, 1.4, 1.4),
  #       unit = "in")

time_table
```











