---
title: "Histograms of betas"
author: "Josh Alampi"
date: "2025-01-23"
output: html_document
---


# load
```{r}
#clear workspace
rm(list=ls(all=TRUE))

# load packages and functions
source(here::here("libraries.R"))
source(here::here("functions/ggplot functions.R"))

data <- read.csv(here::here("data", "raw", "QR sim study_all models_2025-01-18_1000_sims.csv"))

```

# clean

## Fix labels
```{r}
# Fix method labels
# beta_method_labels <- c("Standard \n(undithered) \nfrequentist", "Dithered \nfrequentist", "Unidthered \nBayesianᵃ")
beta_method_labels <- c("Standard \n(undithered) \nfrequentist", "Dithered \nfrequentist", "Unidthered \nBayesian")

model_labels <- c("M1: Binary X1, iid\nmodel errors", "M2: Binary X1, nid\nmodel errors",
                  "M3: Binary X1, nid\n& heavy-tailed\nmodel errors", "M4: Cnts X1, nid\nmodel errors",
                  "M5: Cnts X1, nid\n& heavy-tailed\nmodel errors")


data <- data %>% 
  # Make method labels more detailed
  mutate(beta_method = case_when(method %in% c("riid_nodith", "xy_nodith") ~ beta_method_labels[1],
                                 method %in% c("riid_dith", "xy_dith") ~ beta_method_labels[2],
                                 method %in% c("bayes_adjv1", "bayes_adjv2", "bayes_unadj") ~ beta_method_labels[3])) %>% 
  ## Set the order
  mutate(beta_method = factor(beta_method, beta_method_labels)) %>% 

  # Make model labels more detailed
  mutate(model = case_when(model == 1 ~ model_labels[1],
                           model == 2 ~ model_labels[2],
                           model == 3 ~ model_labels[3],
                           model == 4 ~ model_labels[4],
                           model == 5 ~ model_labels[5])) %>% 
  
  # Clean up the sample size 'n' labels
  mutate(n = paste0("n = ", n))
```

## Filter
I only need to consider the beta coeficients, NOT the confidence intervals for these figures. Therefore, I only need to include the "Unidthered Bayesian", "Standard (undithered) frequentist", and "Dithered frequentist" methods, but not the sub-methods for calculating their intervals
The "xy_nodith" and "riid_nodith" methods give the exact same beta estimates, so I don't need to include both.
```{r}
data <- data %>% 
  filter(method %in% c("bayes_unadj", "riid_nodith", "riid_dith"))
```

# Make labels
```{r}
xlab <- "Point estimate"
ylab <- "Frequency"

fig_width <- 7.5
fig_height_short <- 5.5
fig_height_long <- 10
```

# Make histograms
## make function
```{r}
make_hist_plot <- function(data, title = "", breaks = waiver(), xlims = c(),
                           text_size_minor = 8, text_size_major = 9, 
                           text_size_title = 10, text_family = "sans",
                           panel.spacing.x = 0.075, panel.spacing.y = 0.05,
                           x_angle = 0, ystrip_angle = 0) {
  fig_hist <- data %>% 
    ggplot() + 
    geom_histogram(aes(x = beta), bins = 50, color="black", fill="grey", width = 0.01) + 
    facet_grid(beta_method ~ model, scales = "free_x", axes = "all_x") +
    geom_vline(aes(xintercept = true_beta), lty = 2) + # add vertical line for the true effect of x1 on y
    theme_bw() +
    # Labels
    labs(title= title, x= xlab, y = ylab) + 
    theme(strip.text.y = element_text(angle = 0)) +
    
    # fix font ----
    theme(text = element_text(family = text_family, colour = "black", size = text_size_minor)) +
    theme(axis.text.x = element_text(size=text_size_minor, colour = "black", angle = x_angle)) + 
    theme(axis.text.y = element_text(size=text_size_minor, colour = "black")) + 
    theme(axis.ticks = element_line(size = 0.5, colour = "black")) + 
    
    theme(axis.title.x = element_text(size=text_size_major, colour = "black")) + 
    theme(axis.title.y = element_text(size=text_size_major, colour = "black")) +
    
    theme(strip.text.x = element_text(size=text_size_major, colour = "black")) +
    theme(strip.text.y = element_text(size=text_size_major, colour = "black", angle = ystrip_angle)) +
    theme(legend.text = element_text(size=text_size_minor, colour = "black")) +
    theme(legend.title = element_text(size=text_size_major, colour = "black")) +
    
    theme(title = element_text(size = text_size_title, colour = "black")) +
    theme(plot.caption = element_text(size = text_size_minor, colour = "grey45", 
                                      hjust = 0)) + # move the caption to the right side 
    
    theme(panel.spacing.y = unit(panel.spacing.y, "inch"),
          panel.spacing.x = unit(panel.spacing.x, "inch") ) +
    scale_x_continuous(breaks = breaks, limits = xlims) 
  
  return(fig_hist)
}
```

## make main fig
```{r}
# Save as a pdf for Environmental Epidemiology
cairo_pdf(filename = paste0("results/ee_hist_main", "_",  Sys.Date(), ".pdf"), 
          width = fig_width, height = fig_height_short, 
          fallback_resolution = 1200) # save as pdf

histograms_tau50 <- data %>% 
  filter(tau == 0.5 & n == "n = 250") %>% 
  make_hist_plot(.)

histograms_tau50

ggsave(file = paste0("results/hist_main", "_",  Sys.Date(), ".png"), 
       width = fig_width, height = fig_height_short, dpi = 600)

```

```{r}
# update function to decrease font size
make_hist_plot <- function(data, title = "", breaks = waiver(), xlims = c(),
                           text_size_minor = 7, text_size_major = 8, 
                           text_size_title = 9, text_family = "sans",
                           panel.spacing.x = 0.0075, panel.spacing.y = 0.005,
                           x_angle = 0, ystrip_angle = 0){
  fig_hist <- data %>% 
    ggplot() + 
    geom_histogram(aes(x = beta), bins = 50, color="black", fill="grey", width = 0.01) + 
    facet_grid(beta_method ~ model, scales = "free_x", axes = "all_x") +
    geom_vline(aes(xintercept = true_beta), lty = 2) + # add vertical line for the true effect of x1 on y
    theme_bw() +
    # Labels
    labs(title= title, x= xlab, y = ylab) + 
    theme(strip.text.y = element_text(angle = 0)) +
    
    # fix font ----
    theme(text = element_text(family = text_family, colour = "black", size = text_size_minor)) +
    theme(axis.text.x = element_text(size=text_size_minor, colour = "black", angle = x_angle)) + 
    theme(axis.text.y = element_text(size=text_size_minor, colour = "black")) + 
    theme(axis.ticks = element_line(size = 0.5, colour = "black")) + 
    
    theme(axis.title.x = element_text(size=text_size_major, colour = "black")) + 
    theme(axis.title.y = element_text(size=text_size_major, colour = "black")) +
    
    theme(strip.text.x = element_text(size=text_size_major, colour = "black")) +
    theme(strip.text.y = element_text(size=text_size_major, colour = "black", angle = ystrip_angle)) +
    theme(legend.text = element_text(size=text_size_minor, colour = "black")) +
    theme(legend.title = element_text(size=text_size_major, colour = "black")) +
    
    theme(title = element_text(size = text_size_title, colour = "black")) +
    theme(plot.caption = element_text(size = text_size_minor, colour = "grey45", 
                                      hjust = 0)) + # move the caption to the right side 
    
    theme(panel.spacing.y = unit(panel.spacing.y, "inch"),
          panel.spacing.x = unit(panel.spacing.x, "inch") ) +
    scale_x_continuous(breaks = breaks, limits = xlims) 
  
  return(fig_hist)
}

title_a <- "A) n = 100, τ = 0.1"
title_b <- "B) n = 100, τ = 0.5"
title_c <- "C) n = 100, τ = 0.9"

title_d <- "D) n = 250, τ = 0.1"
title_e <- "E) n = 250, τ = 0.5"
title_f <- "F) n = 250, τ = 0.9"

title_g <- "G) n = 500, τ = 0.1"
title_h <- "H) n = 500, τ = 0.5"
title_i <- "I) n = 500, τ = 0.9"

data_100 <- filter(data, n == "n = 100")
data_250 <- filter(data, n == "n = 250")
data_500 <- filter(data, n == "n = 500")

hist_abc <- ggarrange(data_100 %>% 
                        filter(tau == 0.1) %>% 
                        make_hist_plot(., title_a),
                      data_100 %>% 
                        filter(tau == 0.5) %>% 
                        make_hist_plot(., title_b),
                      data_100 %>% 
                        filter(tau == 0.9) %>% 
                        make_hist_plot(., title_c),
                      nrow = 3)  
ggsave(file = paste0("results/hist_n100", "_",  Sys.Date(), ".png"), 
       width = fig_width, height = fig_height_long, dpi = 600)

hist_def <- ggarrange(data_250 %>% 
                        filter(tau == 0.1) %>% 
                        make_hist_plot(., title_d),
                      data_250 %>% 
                        filter(tau == 0.5) %>% 
                        make_hist_plot(., title_e),
                      data_250 %>% 
                        filter(tau == 0.9) %>% 
                        make_hist_plot(., title_f),
                      nrow = 3)  
ggsave(file = paste0("results/hist_n250", "_",  Sys.Date(), ".png"), 
       width = fig_width, height = fig_height_long, dpi = 600)

hist_ghi <- ggarrange(data_500 %>% 
                        filter(tau == 0.1) %>% 
                        make_hist_plot(., title_g),
                      data_500 %>% 
                        filter(tau == 0.5) %>% 
                        make_hist_plot(., title_h),
                      data_500 %>% 
                        filter(tau == 0.9) %>% 
                        make_hist_plot(., title_i),
                      nrow = 3)  
ggsave(file = paste0("results/hist_n500", "_",  Sys.Date(), ".png"), 
       width = fig_width, height = fig_height_long, dpi = 600)

hist_abc

```

```{r}

```

