---
title: "Performance figs 2025_v2"
output: html_document
date: "2025-02-26"
---


# load
```{r}
#clear workspace
rm(list=ls(all=TRUE))

# load packages and functions
source(here::here("libraries.R"))
source("functions/ggplot functions.R") # To save time, I created a function "gg_geom_point()" which I will call to create the figures. 

res <- read.csv("data/cleaned/cleaned SS_all models_2025-01-18_1000_sims.csv") %>% 
  mutate(tau = as.factor(tau)) %>% 
  mutate(blank = F) # add a blank row, used to edit the shape of figures
```


# clean
## Fix labels
```{r}
# Fix method labels
# beta_method_labels <- c("Standard \n(undithered) \nfrequentist", "Dithered \nfrequentist", "Unidthered \nBayesianᵃ")
beta_method_labels <- c("Standard \n(undithered) \nfrequentist", "Dithered \nfrequentist", "Unidthered \nBayesian")

# beta_method_labels_ref <- c("Standard \n(undithered) \nfrequentist \n[reference]", "Dithered \nfrequentist", "Unidthered \nBayesianᵃ") # Label reference group
beta_method_labels_ref <- c("Standard \n(undithered) \nfrequentist \n[reference]", "Dithered \nfrequentist", "Unidthered \nBayesian") # Label reference group

# beta_method_labels_int <- c("Standard \n(undithered) \nfrequentist", "Dithered \nfrequentist", "Unidthered \nBayesianᶜ")
beta_method_labels_int <- c("Standard \n(undithered) \nfrequentist", "Dithered \nfrequentist", "Unidthered \nBayesian")

model_labels <- c("M1: Binary X1, iid\nmodel errors", "M2: Binary X1, nid\nmodel errors",
                  "M3: Binary X1, nid\n& heavy-tailed\nmodel errors", "M4: Cnts X1, nid\nmodel errors",
                  "M5: Cnts X1, nid\n& heavy-tailed\nmodel errors")

# int_method_labels <- c("Rank (not dith.)ᵃ", "Rank (dith.)ᵃ", 
#                       "xy (not dith.)ᵇ", "xy (dith.)",
#                       "Bayes. (unadj.)ᶜ", "Bayes. (adj)ᶜ")
# 
# int_method_labels_cov <- c("Rank (not\ndith.)ᵃ", "Rank\n(dith.)ᵃ", 
#                           "xy (not\ndith.)ᵇ", "xy\n(dith.)",
#                           "Bayes.\n(unadj.)ᶜ", "Bayes.\n(adj)ᶜ")

int_method_labels <- c("Rank (not dith.)", "Rank (dith.)", 
                      "xy-bootstrap (not dith.)", "xy-bootstrap (dith.)",
                      "Bayesian (unadj.)", "Bayesian (adj.)")

int_method_labels_cov <- c("Rank (not\ndithered)", "Rank\n(dithered)", 
                          "xy-bootstrap\n(not\ndithered)", "xy-bootstrap\n(dithered)",
                          "Bayesian\n(unadjusted)", "Bayesian\n(adjusted)")

res <- res %>% 
  # Set the order of all methods
  # mutate(method = factor(method, c("riid_nodith", "riid_dith", 
  #                                  "xy_nodith", "xy_dith",
  #                                  "bayes_unadj", "bayes_adjv2"))) %>% 

  # Add a new column "beta_method", which groups methods together
  mutate(beta_method = case_when(method %in% c("riid_nodith", "xy_nodith") ~ beta_method_labels[1],
                                 method %in% c("riid_dith", "xy_dith") ~ beta_method_labels[2],
                                 method %in% c("bayes_adjv1", 
                                               "bayes_adjv2", "bayes_unadj") ~ beta_method_labels[3])) %>% 
  ## Set the order
  mutate(beta_method = factor(beta_method, beta_method_labels)) %>% 
  
  
  # Add a new column "beta_method_ref", which is the same as above, but labels the reference group
  mutate(beta_method_ref = case_when(beta_method == beta_method_labels[1] ~ beta_method_labels_ref[1],
                                     .default = beta_method)) %>% 
  ## Set the order
  mutate(beta_method_ref = factor(beta_method_ref, beta_method_labels_ref)) %>%
  
  
  # Add a new column "beta_method_int", which changes the superscripts for the 'interval' figures
  mutate(beta_method_int = case_when(beta_method == beta_method_labels[3] ~ beta_method_labels_int[3],
                                     .default = beta_method)) %>% 
  ## Set the order
  mutate(beta_method_int = factor(beta_method_int, beta_method_labels_int)) %>%
  
  
  # Fix method labels
  mutate(int_method = case_when(method == "riid_nodith" ~ int_method_labels[1],
                                method == "riid_dith" ~ int_method_labels[2],
                                method == "xy_nodith" ~ int_method_labels[3],
                                method == "xy_dith" ~ int_method_labels[4],
                                method == "bayes_unadj" ~ int_method_labels[5],
                                method == "bayes_adjv2" ~ int_method_labels[6]
                                )) %>% 
  mutate(int_method_cov = case_when(method == "riid_nodith" ~ int_method_labels_cov[1],
                                    method == "riid_dith" ~ int_method_labels_cov[2],
                                    method == "xy_nodith" ~ int_method_labels_cov[3],
                                    method == "xy_dith" ~ int_method_labels_cov[4],
                                    method == "bayes_unadj" ~ int_method_labels_cov[5],
                                    method == "bayes_adjv2" ~ int_method_labels_cov[6]
                                    )) %>% 
  ## Set the order
  mutate(int_method = factor(int_method, int_method_labels)) %>%
  mutate(int_method_cov = factor(int_method_cov, int_method_labels_cov)) %>%

  # Make model labels more detailed
  mutate(model = case_when(model == 1 ~ model_labels[1],
                           model == 2 ~ model_labels[2],
                           model == 3 ~ model_labels[3],
                           model == 4 ~ model_labels[4],
                           model == 5 ~ model_labels[5]))

```


## filter
I only need to consider the beta coeficients, NOT the confidence intervals for bias and empSE. Therefore, I only need to include the "Unidthered Bayesian", "Standard (undithered) frequentist", and "Dithered frequentist" methods, but not the sub-methods for calculating their intervals
The "xy_nodith" and "riid_nodith" methods give the exact same beta estimates, so I don't need to include both.
```{r}
res_short <- res %>%
  filter(method %in% c("bayes_unadj", "riid_nodith", "riid_dith"))
```


# Prep figures

## labels
```{r}
# y_cov <- "Empirical covarage probabilities (95% Monte Carlo intervals)"
# y_intwidthvar <- "Empirical standard deviation of 90% interval widths (95% Monte Carlo intervals)"
# y_bias <- "Bias (95% Monte Carlo intervals)"
# y_empSE <- "Empirical standard error of point estimates (95% Monte Carlo intervals)"
# y_empSE_rel <- "Relative % difference in empirical standard error of point estimates (95% Monte Carlo intervals)"
# y_blank <- ""

y_cov <- "Empirical covarage probabilities (±1 Monte Carlo standard error)"
y_intwidthvar <- "Empirical standard deviation of 90% interval widths (±1 Monte Carlo SE)"
y_bias <- "Bias (±1 Monte Carlo standard error)"
y_empSE <- "Empirical standard error of point estimates (±1 Monte Carlo standard error)"
y_empSE_rel <- "Relative (%) difference in the precision of point estimates (±1 Monte Carlo SE)"
y_blank <- ""

x_lab <- "Quantile (τ)"
col_title <- "Method"

fig_width <- 7.5
fig_height_short <- 5.5
fig_height <- 8.5
fig_height_long <- 10
```

<!-- ## colours -->
<!-- ```{r} -->
<!-- show_col(hue_pal()(3)) -->
<!-- # col_3 <- c("#7CAE00", "#00BFC4", "#C77CFF") -->
<!-- ``` -->

# Make figures
## bias
```{r}
make_bias_plot <- function(res, title = NULL) {
  fig_bias <- res %>%
    # Use function 'gg_geom_point' to set defaults and ensure consistency
    gg_geom_point(x = tau, y = (bias), 
                  ymin = (bias)-(bias_se), 
                  ymax = (bias) + (bias_se),
                  p_dodge = T, 
                  facet_type = "grid", hori = "model", vert = "n", scales = "free_x",
                  
                  shape = blank, shape_vec = c(1), shape_title = NULL,
                  colour = beta_method, col_title = col_title,
                  legend.direction = "horizontal",
                  
                  title = title, ylab = y_bias, xlab = x_lab,
                  coord_flip = F, ystrip_angle = 0
                  # panel.spacing.x = 0.2, panel.spacing.y = 0.05
                  ) + 
    scale_y_continuous(labels = number_format(accuracy = 0.1) ) +
    theme(legend.box.spacing = unit(0.0001, units = "inch"))
  
  return(fig_bias)
}

# Save as a pdf for Environmental Epidemiology
cairo_pdf(filename = paste0("results/ee2_fig_bias_all_n", "_",  Sys.Date(), ".pdf"), 
          width = fig_width, height = fig_height_short, 
          fallback_resolution = 1200) # save as pdf

fig_bias_all <- make_bias_plot(res_short) 
fig_bias_all

ggsave(file = paste0("results/fig_bias_all_n", "_",  Sys.Date(), ".png"),
       width = fig_width, height = fig_height_short, dpi = 600)

```


<!-- ## empSE -->
<!-- ```{r} -->
<!-- make_empSE_plot <- function(res, title = NULL) { -->
<!--   plot <- res %>%  -->
<!--     # Use function 'gg_geom_point' to set defaults and ensure consistency for all figures -->
<!--     gg_geom_point(x = tau, y = empSE,  -->
<!--                   ymin = empSE-(empSE_se), ymax = empSE + (empSE_se), -->
<!--                   shape = blank, shape_vec = c(1), shape_title = NULL, -->
<!--                   colour = beta_method, col_title = col_title,  -->
<!--                   p_dodge = T, -->
<!--                   facet_type = "grid", hori = "model", vert = "n", -->
<!--                   legend.direction = "horizontal", -->
<!--                   scales = "free_x", -->
<!--                   title = title, ylab = y_empSE, xlab = x_lab, -->
<!--                   yint = NULL, ystrip_angle = 0, -->
<!--                   # panel.spacing.x = 0.2, panel.spacing.y = 0.05 -->
<!--                   ) +  -->
<!--     scale_y_continuous(labels = number_format(accuracy = 0.01) ) + -->
<!--     theme(legend.box.spacing = unit(0.0001, units = "inch")) -->

<!--   return(plot) -->
<!-- } -->

<!-- fig_empSE_all <- make_empSE_plot(res_short) -->
<!-- fig_empSE_all -->

<!-- ggsave(file = paste0("results/fig_empSE_all_n", "_",  Sys.Date(), ".png"), -->
<!--        width = fig_width, height = fig_height_short, dpi = 600) -->
<!-- ``` -->

## empSE_rel
```{r}
make_empSE_rel_plot <- function(res, title = NULL) {
  plot <- res %>% 
    # Use function 'gg_geom_point' to set defaults and ensure consistency for all figures
    gg_geom_point(x = tau, y = empSE_rel, 
                  ymin = empSE_rel-(empSE_rel_se), ymax = empSE_rel + (empSE_rel_se),
                  shape = blank, shape_vec = c(1), shape_title = NULL,
                  colour = beta_method_ref, col_title = col_title, 
                  p_dodge = T, 
                  facet_type = "grid", hori = "model",  vert = "n",
                  legend.direction = "horizontal",
                  scales = "free_x",
                  title = title, ylab = y_empSE_rel, xlab = x_lab,
                  yint = 0, ystrip_angle = 0,
                  # panel.spacing.x = 0.2, panel.spacing.y = 0.05
                  )  +
    theme(legend.box.spacing = unit(0.0001, units = "inch"))
  
  return(plot)
}

# Save as a pdf for Environmental Epidemiology
cairo_pdf(filename = paste0("results/ee3_fig_empSE_rel_all_n", "_",  Sys.Date(), ".pdf"), 
          width = fig_width, height = fig_height_short, 
          fallback_resolution = 1200) # save as pdf

fig_empSE_rel_all <- make_empSE_rel_plot(res_short)
fig_empSE_rel_all

ggsave(file = paste0("results/fig_empSE_rel_all_n", "_",  Sys.Date(), ".png"),
       width = fig_width, height = fig_height_short, dpi = 600)
```


## cov
```{r}
make_cov_plot <- function(res, title = NULL, cov_limit = 0.65) {
  cov_limit <- cov_limit
  
  data_tmp <- res %>% 
    # if coverage rate is out of bounds, then it should be replaced with an arrow
    filter(is.na(cov) == F) %>% # Undith xy was not run for models 4 and 5. These values need to be removed
    mutate(alpha = ifelse(cov < cov_limit , 1, 0)) %>% 
    mutate(cov_se = ifelse(cov < cov_limit, 0, cov_se)) %>%
    mutate(cov = ifelse(cov < cov_limit, 0, cov)) 
    
  # Use function 'gg_geom_point' to set defaults and ensure consistency
  plot <- gg_geom_point(.data = data_tmp, x =tau, y = cov, ymin = cov - (cov_se), ymax = cov + (cov_se),
                        shape = blank, shape_vec = c(1), shape_title = NULL,
                        colour = beta_method_int, col_title = col_title, 
                        legend.direction = "horizontal",
                        facet_type = "grid", hori = "model", vert = "int_method_cov",  p_dodge = F,
                        ylab = y_cov, xlab = x_lab, title = title,
                        text_size_minor = 8, text_size_major = 10, yint = 0.9, 
                        limits = c(cov_limit,1), coord_flip = F, ystrip_angle = 0,
                        # panel.spacing.x = 0.15, panel.spacing.y = 0.05
                        ) + 
    scale_y_continuous(labels = scales::number_format(accuracy = 0.1) ) +
    theme(legend.box.spacing = unit(0.0001, units = "inch")) +
    
    # Add an arrow to denote out of range values. This arrow will have an alpha=0 (i.e. be invisible) when 'cov' value is within range.
    geom_segment(data = data_tmp, 
                 aes(y = 0.8, yend = cov_limit, 
                     x = tau, xend = tau, alpha = alpha),
                 arrow = arrow(length = unit(0.05, "inch"))) +
    scale_alpha(range = c(0, 1)) + 
    facet_grid(int_method_cov ~ model) + 
    guides(alpha = "none") 
  
  return(plot)
}

# Save as a pdf for Environmental Epidemiology
cairo_pdf(filename = paste0("results/ee4_fig_cov_main_250", "_",  Sys.Date(), ".pdf"), 
          width = fig_width, height = fig_height_short, 
          fallback_resolution = 1200) # save as pdf

fig_cov_250_main <- res %>% #remake without a title (this plot will go in the manuscript)
  filter(n == "n = 250") %>% 
  make_cov_plot(.)
fig_cov_250_main

ggsave(file = paste0("results/fig_cov_main_250", "_",  Sys.Date(), ".png"),
       width = fig_width, height = fig_height, dpi = 600)

# Make plots for the appendix (remainint sample sizes)
title_a <- "A) n = 100"
title_b <- "B) n = 500"
# title_c <- "C) n = 500"

fig_cov_100 <- res %>% 
  filter(n == "n = 100") %>% 
  make_cov_plot(., title_a)

fig_cov_100
ggsave(file = paste0("results/fig_cov_100", "_",  Sys.Date(), ".png"),
       width = fig_width, height = fig_height, dpi = 600)

# fig_cov_250 <- res %>% 
#   filter(n == "n = 250") %>% 
#   make_cov_plot(., title_b)
# fig_cov_250
# 
# ggsave(file = paste0("results/fig_cov_250", "_",  Sys.Date(), ".png"),
#        width = fig_width, height = fig_height, dpi = 600)

fig_cov_500 <- res %>% 
  filter(n == "n = 500") %>% 
  make_cov_plot(., title_b)

fig_cov_500
ggsave(file = paste0("results/fig_cov_500", "_",  Sys.Date(), ".png"),
       width = fig_width, height = fig_height, dpi = 600)

```


## intwidth var fig
Main plot: tau = 0.9 only. Remake 3 more versions with other taus
```{r}
make_intwidthEmpSE_plot <- function(res, title = NULL, ylab = "") {
  plot <- res %>% 
    filter(intwidth_empSE != 0 ) %>% # Undith xy was not run for models 4 and 5. It was given variance ==0, need to remove
    # Use function 'gg_geom_point' to set defaults and ensure consistency for all figures
    gg_geom_point(x = int_method, y = intwidth_empSE, 
                  ymin = intwidth_empSE - (intwidth_empSE_se), 
                  ymax = intwidth_empSE + (intwidth_empSE_se),
                  shape = blank, shape_vec = c(1), shape_title = NULL,
                  colour = beta_method_int, col_title = col_title, 
                  p_dodge = F,
                  legend.direction = "horizontal",
                  facet_type = "grid", hori = "model", vert = "n", scales = "free",
                  ylab = ylab, title = title,
                  xlab = "Quantile regression interval method",
                  text_size_minor = 8, text_size_major = 10,
                  yint = NULL, coord_flip = F, ystrip_angle = 0,
                  panel.spacing.x = 0.1, panel.spacing.y = 0.05) + 
    scale_y_continuous(labels = number_format(accuracy = 0.1) ) +
    theme(axis.text.x = element_text(angle = 315, hjust =0, vjust = 0.5)) +
    theme(legend.box.spacing = unit(0.0001, units = "inch"))
  
  return(plot)
}

# Save as a pdf for Environmental Epidemiology
cairo_pdf(filename = paste0("results/ee5_fig_intwidthEmpSE_main_tau_0.9", "_",  Sys.Date(), ".pdf"), 
          width = fig_width, height = fig_height_short, 
          fallback_resolution = 1200) # save as pdf

# Make main plot (tau =0.9 only)
fig_intwidthvar_main <- res %>% 
  filter(tau == 0.9) %>% 
  make_intwidthEmpSE_plot(., ylab = y_intwidthvar)

fig_intwidthvar_main

ggsave(file = paste0("results/fig_intwidthEmpSE_main_tau_0.9", "_",  Sys.Date(), ".png"),
       width = fig_width, height = fig_height_short, dpi = 600)


# Make plots for the appendix (remaining taus)
# fix labels
title_a <- "A) τ = 0.1"
title_b <- "B) τ = 0.5"
# title_c <- "C) τ = 0.9"

# make figs
fig_intwidthvar_a <- res %>%
  filter(tau == 0.1) %>%
  make_intwidthEmpSE_plot(., title_a, y_intwidthvar)

ggsave(file = paste0("results/fig_intwidthEmpSE_a", "_",  Sys.Date(), ".png"),
       width = fig_width, height = fig_height_short, dpi = 600)



fig_intwidthvar_b <- res %>%
  filter(tau == 0.5) %>%
  make_intwidthEmpSE_plot(., title_b, y_intwidthvar)

ggsave(file = paste0("results/fig_intwidthEmpSE_b", "_",  Sys.Date(), ".png"),
       width = fig_width, height = fig_height_short, dpi = 600)



# fig_intwidthvar_a_b <- ggarrange(res %>% ############## CUT. Easier to present this figure on different pages
#                                    filter(tau == 0.1) %>%
#                                    make_intwidthEmpSE_plot(., title_a),
#                                  res %>%
#                                    filter(tau == 0.5) %>%
#                                    make_intwidthEmpSE_plot(., title_b),
#                                 
#                                  nrow = 2, common.legend = T, legend = "bottom") %>%
#   annotate_figure(left = text_grob(y_intwidthvar, rot = 90, size = 9, vjust = 3))
# 
# ggsave(file = paste0("results/fig_intwidthEmpSE_a_b", "_",  Sys.Date(), ".png"),
#        width = fig_width, height = fig_height_long, dpi = 600)
# 
# fig_intwidthvar_c <- res %>%
#   filter(tau == 0.9) %>%
#   make_intwidthEmpSE_plot(., title_c, y_intwidthvar)
```



```{r}

```


